Put Title here

### 2018-09-11 General WLD combinatorics.tex
tidied up and edited section 2

\section{Wilson Loop diagrams as matroids}

In \cite{wilsonloop}, Amat and Agarwala show that admissible Wilson loop diagrams with $k$ propagators corresponds to positroids of rank $k$ (a matroid that can be represented by an element of $\Gr(k, n)$.)

\begin{thm}[\cite{wilsonloop}] An admissible Wilson loop diagram, $W =(\cP, n)$, defines a matroid with base set $[n]$. The independent sets are exactly those subsets $V \subseteq [n]$ such that $\not \exists U\subseteq V$ satisfying $|\Prop(U)|> |U|$. \label{thm:WLDmatroid}\end{thm}

This does not, however, show that the matroid associated to each Wilson loop diagram is unique. In fact, in \cite{wilsonloop}, the authors show that if two diagrams are equivalent then they define the same matroid [cite theorem here].

In the remainder of this section, we show that two admissible Wilson loop diagrams define the same matroid if an only if they are equivalent. Furthermore, we count the number of Wilson loop diagrams in each equivalence class. The first of these are done by considering Wilson loop diagrams as matroids. The second, by considering polygon partitions of Wilson Loop Diagrams.

To this end, we recall some definitions and go through a series of new intermediary steps.

First we recall a few facts about matroids, with particulars about those defined by Wilson loop diagrams.

\begin{dfn} For a Wilson Loop diagram $W = (\cP, [n])$, the associated matroid $M(W)$ can be defined by any of the following sets of subsets
\begin{enumerate}
\item An set $V \subseteq [n]$ is independent if there is no subset that supports fewer propagators than vertices of that subset: \bas \not \exists U \subseteq V \; \textrm{s.t. } |\Prop (U)| < |U|\;. \eas A basis of $W$ is any independent set, $B \subset [n]$ of size $|B| = |\cP|$.
\item The rank of a set $|V| \subset [n]$, is bounded above by $\rk(V) \leq \min\{ |V| , |\Prop(V)|\}$. If $V$ is an independent set, then $\rk(V) = |V|$. Furthermore, if $v \in [n]$, and $q \in \Prop(v)$, then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, $\rk(V \cup v) = \rk(V) +1$. That is, adding a vertex that supports a new propagator to a set increases the rank of the set.
\item A circuit of a matroid is any set $C \subset [n]$ such that every proper subset $U \subsetneq C$ is independent. That is, $\rk(C) = |C|-1$. The union of circuits is called a cycle. If $C$ is a circuit of an admissible Wilson loop diagram, and $e \in C$, \bas |\Prop(C\setminus e))| \leq |C\setminus e| = \rk(C) \leq \min\{|C|, |\Prop(C\setminus e)|\}\;.\eas
    In other words, if $C$ is a circuit of a Wilson loop diagram, then \bas \rk(C) = |C|-1 = |\Prop(C)| \;. \eas
\item A flat, $F$, of a matroid is a set that is maximal for its rank. That is, for all $e \not \in F$ $\rk(F \cup e) = \rk(F) +1$. If $W$ is an admissible Wilson loop diagram, any group of propagators defines a flat, $F(P) = V(P^c)^c$, called a propagator flat. To see that this is flat, let $v \in V(P^c)$. We have that there is a $q \not \in P$ such that $q \in \Prop(v)$. Therefore, by the rank function, $\rk( F(P) \cup v))> \rk(F(P))$. Furthermore, in \cite{wilsonloops}, the authors show that if $F$ is a cyclic flat, then there is a $P \subset \cP$ such that $F = F(P)$.

\end{enumerate}
\end{dfn}

Note that one may substitute values into $C(W)$ to realize the matroid assoicated to $W$ in this manner.

It is worth noting that it is not necessary to list all the flats in a matroid to defined it completely, only the cyclic flats, and those that are independent sets. Namely, if $F$ is a dependent flat, it contains a circuit. Let $C \subset F$ be the largest cycle contained in $F$. Then two things are true.

\begin{lem} Let $W$ be a Wilson loop diagram, and $F$ is a flat. Let $C \subset F$ be the largest cycle contained in $F$. Then the following is true:
\begin{enumerate}
\item $C = F(\Prop (C))$ is a cyclic flat.
\item $F \setminus C$ is an independent set. If $W$ does not contain any vertices of rank $0$, that is, all vertices support at least $1$ propagator, then $F \setminus C$ is an independent flat.
\end{enumerate}
\end{lem}

\begin{proof}
If $F$ is an independent flat, then $C = \emptyset$ and the statement is trivially true.

If $F$ is a dependent set, $F$ contains a circuit.

To see the first point, note that $\rk(C) = |\Prop(C)| \geq F(\Prop (C)|$.  For any $v\in C$ we have that $\Prop(v) \subset \Prop(C)$. This implies that $v \in F(\Prop(C))$. Therefore $C \subset F(\Prop(C))$. This implies that $|\Prop(C)| = F(\Prop (C)|$. Suppose there is a $v \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank. The set $B \cup v$ is dependent and therefore contains a circuit. Therefore, $C \cup b$ is a cycle, which is a contradiction to $C$ being the maximal cycle in $F$.

To see the second point, note that $F \setminus C$ is independent (as it contains no circuits). Furthermore, it is a flat. For any $e \not \in F$, $\rk (F \setminus C \cup e) = \rk (F\setminus C) +1$, since $F$ is a flat. For any $e \in C$,  if $\rk (F \setminus C \cup e) = \rk (F\setminus C)$, the set $F \setminus C \cup e$ is a dependent set. This occures either if $\rk{e} = 0$, or $F\setminus C \cup e$ contains a circuit of at least $2$ elements, which contradicts the hypothesis that $C$ is the largest cycle in $F$.
\end{proof}

Therefore, in the sequel, one only needs worry about cyclic and independent flats.

\begin{cor} \label{classifyflats}
If $F$ is a flat of a Wilson loop diagram, it can be written as the union of a cyclic propagator flat and an independent flat. \end{cor}

@@

Finally, we note the differences between subdiagrams of Wilson loop diagrams, and restrictions or cotractions of the associated matroids.

Given any matroid, one may restrict it to a subset of the base set. The bases of the restricted matroid come from intersecting bases of the original with the subset. It is worth noting that a the matroid defined by a subdiagram is different from the restriction of the matroid of a Wilson loop diagram to a set of vertices.

\begin{dfn} \label{restrictiondfn}
For $W = (\cP, n)$, the restricted diagram, $W|_V$ is the matroid defined by only looking at the vertices $V \subset [n]$.
\end{dfn}

The key difference between a subdiagram and a restriction is that the propagator support function does not change in the case of restriction, while it may in the case of a subdiagram. In particular, for $v \in V$, $\Prop_W(v) = \Prop_{W|_V}(v)$, while $\Prop_{(P, V(P))} (v) = \Prop_W(v) \cap P$.

A subdiagram is more closely related to a contracted matroid.

\begin{lem} \label{contractsubdiaglem}
For $W = (\cP, [n])$, consider a subdiagram $(P^c, V(P^c))$. The complementary set of vertices is a propagator flat: $F(P) = V(P^c)^c$. If this has maximal rank, $\rk F(P) = |P|$, then the subdiagram defines the same matroid as is formed by contracting the original matroid by $F(P)$: \bas M((P^c, V(P^c))) = M(W)/ F(P) \;. \eas
\end{lem}

In the sequel, where it doesn't cause confusion, I conflate a Wilson loop diagram with the matroid it defines.

In order to study exact Wilson loop diagrams, we further introduce the notion of a polygon partition of $W$.

\begin{dfn}
  Let $W = (\cP, n)$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denotes $\tau(W)$ is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$ for each propagator of $W$.  If $p= (i,j) \in \cP$, it defines a chord connecting the verices $i$ and $j$
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, $\tau(W)$ is a simple planar graph wihose outer face is a cycle. It is embedded such that the vertices all like on this infinite face. These vertices are cyclically ordered, with a coice of first vertex, giving it an additional compatible linear order.
\end{lem}

>>>>

\section{Equivalence classes of Wilson loop diagrams}

In \cite{wilsonloop}, Agarwala and Amat show that Wilson loop diagrams can be interpreted as positroids, a certain well behaved class of realizable matroids (this correspondence is stated precisely in Theorem \ref{thm WLD defines matroid} below). This opens up the study of Wilson loop diagrams to techniques from geometry and combinatorics.

[outline of section, postponed until section structure is finalised]

The main results of this section are the following: we show that two admissible Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}), and we obtain a formula for the number of admissible Wilson loop diagrams in each equivalence class (Corollary \ref{number of equiv diagrams}).  [sentence about why this matters]

\subsection{Wilson loop diagrams as matroids}\label{sec matroid background}

We first give a quick summary of the matroid terminology that we will need; it is not intended as a comprehensive introduction to matroids and the interested reader is referred to [ref].

A {\em matroid} $M = (E,\cB)$ consists of a finite ground set $E$ and a non-empty family $\cB \subseteq \cP(E)$ whose elements satisfy the {\em basis exchange property}: for any distinct $B_1,B_2 \in \cB$ and any $a \in B_1 \setminus B_2$, there exists some $b \in B_2 \setminus B_1$ such that $(B_1 \setminus \{a\})\cup \{b\} \in \cB$ as well. The elements of $\cB$ are called the {\em bases} of the matroid. Note that the basis exchange property immediately implies that all bases have the same size.

A subset $A \subseteq E$ is called {\em independent} in $M$ if $A \subseteq B$ for some $B \in \cB$, and {\em dependent} else. The {\em rank}  $\rk(A)$ of a subset $A \subseteq E$ is the size of the largest independent set contained in $A$. The rank of the matroid itself is defined to be $\rk(E)$.

A {\em circuit} in $M$ is a minimally dependent set. That is, it is a set $C \subseteq E$ such that $C$ is dependent but $C \setminus \{e\}$ is independent for any $e \in C$. A union of circuits is called a {\em cycle}. On the other hand, a {\em flat} is a maximally dependent set, i.e. a set $F \subseteq E$ such that $\rk(F \cup \{e\}) = \rk(F) + 1$ for any $e \in E \setminus F$. Unsurprisingly, a {\em cyclic flat} is a set which is both a flat and a cycle. The set of circuits in a matroid uniquely defines that matroid, as does the set of flats; thus one could specify a matroid by listing its independent sets, bases, circuits, or flats. [ref]

Finally, we describe several important types of matroid. A matroid of rank $k$ with a ground set of size $n$ is called {\em realizable} if there exists some $A \in Gr(k,n)$ whose non-zero $k\times k$ minors are exactly those with columns indexed by elements of $\cB$. A {\em positroid} is a matroid which can be realized by an element of the totally nonnegative Grassmannian $\Gr(k,n)$. Finally, a {\em uniform matroid} of rank $r$ is a matroid in which any set of size $\leq r$ is independent.

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite{wilsonloop}, Agarwala and Amat show that every admissible Wilson loop diagram with $k$ propagators defines a positroid of rank $k$, and that the independent sets can be read directly from the diagram:

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} Any admissible Wilson loop diagram $W =(\cP, n)$ defines a matroid $M(W)$ with ground set $[n]$. The independent sets are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \label{thm:WLDmatroid}\end{thm}
In other words, the independent sets of $M(W)$ correspond to the sets of vertices in $W$ such that no subset supports fewer propagators than the vertices it contains.

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note that since vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$, $M(W)$ can also be thought of as the matroid realized by $C(W)$.

% The identification of Theorem \ref{thm WLD defines matroid} is not 1-1: by \cite[Theorem 1.18]{wilsonloop}, if two admissible Wilson loop diagrams are equivalent (as in Definition~\ref{equivdfn}) then they define the same matroid.

Let $W = (\cP,n)$ be an admissible Wilson loop diagram, and $M(W)$ its associated matroid. Where it will not cause confusion we conflate the two objects, identifying vertices of $W$ with elements of the ground set $[n]$ in $M(W)$.

In particular, this allows us to prove results about $M(W)$ by considering the behavior of propagators in $W$. We record a few elementary facts about the rank and cycles of $M(W)$ here as an example of this.\note{It would be nice to have a few more; do we need any others, and/or are there any cool ones?}

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}.
\end{proof}

The equivalence relation on Wilson loop diagrams is defined in terms of exact subdiagrams; thus in order to understand the equivalence, we need a way to extract and compare exact subdiagrams. We do this via the notion of a polygon partition of $W$.

\begin{dfn}
  Let $W = (\cP, n)$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, n)$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices. Furthermore, since the outer polygon of $W$ is a cycle, the outer face of $\tau(W)$ is also a cycle. Since $W$ is admissible, no pairs of propagators cross. Therefore, it is a planar embedding. Similarly, $W$ does not admit any propagators of the form $p = (i, i+1)$; therefore there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Finally, there does not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$.
 \end{proof}

\begin{eg}\label{WLDtopolygonpartition}
In this example we return to two of the Wilson loop diagrams in Example \ref{eg:equivdiags}. We can pair diagrams with their polygon partitions as follows:
\bas W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad; \quad
\tau(W_1) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{5}{8}
    \drawchord{5}{7}
	\end{scope}
	\end{tikzpicture}
\eas and
\bas W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
    \drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}	
		\end{scope}
	\end{tikzpicture}\quad; \quad
\tau(W_3) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
	\end{scope}
	\end{tikzpicture} .
\eas

Recall that a planar embedding of a graph is a \emph{triangulation} if all faces, except possibly the infinite face, are triangles.

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

\begin{dfn}
 A {\em decomposition} of a polygon partition $\tau(W)$ is a set of 2-connected induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

\begin{eg} \label{eg: unique decomposition} For the Wilson loop diagrams and polygon partitions in Example \ref{WLDtopolygonpartition}, the vertex sets $\{1, 2, 3,4\}$ and $\{5, 6, 7, 8\}$ are maximal triangulated pieces for both $\tau(W_1)$ and $\tau(W_3)$. The vertex set $\{4,5, 8, 1\}$ is not a triangulation in either polygon partition. \note{do we want a picture of a dual tree here? Being lazy}
\end{eg}

Let $T(W)$ be the dual graph of $\tau(W)$ with the vertex corresponding to the infinite face split. In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary. Since $\tau(W)$ is a simple planar graph by Lemma \ref{tausimpleplanarlem}, $T(W)$ is an uniquely defined graph.

We claim that $T(W)$ is a tree. Since $\tau(W)$ is a planar graph with $k$ chord edges and no internal vertices, there are $k+1$ internal faces of $\tau(W)$. Each of these internal faces corresponds to a vertex of $T(W)$, and there are $n$ vertices of $T(W)$ from the splitting of the dual graph, so $T(W)$ has $n+k+1$ vertices in total. On the other hand, $T(W)$ has $n+k$ edges, corresponding to the $n +k$ edges of $\tau(W)$. Therefore, $T(W)$ is a tree.

By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle). The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  Let $f$ be the forest thus obtained. The vertices of $f$ either have valence 1, (if they correspond leaf vertices) or 3,  (if they correspond to triangulated faces). Subtrees with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulations. Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a valence $3$ vertex.

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition. This implies that $v$ has valence greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the decomposition corresponding to $f'$ is not a triangulation. 
\end{proof}

\begin{proof}
Consider any two distinct maximal triangulated pieces of $\tau(W)$. These two pieces correspond to subtrees of $T(W)$ and intersect, at most, at a vertex in the interior of $\tau(W)$. Since the subtrees corresponding to the maximal triangulated pieces are edge disjoint, and the edges of $T(W)$ correspond to the edges of $\tau(W)$, this forces the maximal triangulated pieces to be edge disjoint as well.
\end{proof}

We are now in a position to relate the triangulated pieces of $\tau(W)$ to exact subdiagrams of $W$.

\begin{proof}
Let us first record a few standard facts about polygon triangulations (that is, about triangulations with all vertices on the outer face).  If such a triangulation has $n$ vertices then it has $n$ edges on the polygon (that is, on the outer face) and $n-3$ edges which are not.  No planar graph with the same vertices and the same outer face can have more edges than the triangulation, and every such simple graph with $n-3$ edges off the outer face is a triangulation.

Since $W$ is admissible, by Lemma \ref{tausimpleplanarlem} $\tau(W)$ is a simple graph. Furthermore, $t$ is not all of $\tau(W)$.

Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decompositionlem}; note that $t$ cannot be equal to $\tau(W)$ by the definition of admissible diagrams.

If $t$ has 2 vertices, then $t$ has two connected components. Therefore, $t$ corresponds to a propagator that connects two non-adjacent edges. Therefore, the trivial triangulation is a trivial exact subdiagram.

Now suppose that $t$ has $m>2$ vertices; we count how many edges of $t$ are not on the outer face of $\tau(W)$.  Consider the intersection of $t$ with the outer face of $\tau(W)$: this is a possibly disconnected subgraph of the polygon of $\tau(W)$ with $m$ vertices, call it $S$.  Let $k$ be the number of connected components of $S$.   To join the components $S$ must have $k$ edges not in the outer face of $\tau(W)$.  Furthermore $t$ has $m-3$ edges not in its outer face and so also not in the outer face of $\tau(W)$.  Thus there are $m-3+k$ edges of $t$ not in the outer face of $\tau(W)$.

Each of these $m-3+k$ internal edges correspond to propagators in $W$, call this set $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $k$ connected compnents of the outer polygon of $W$. Thus the set $V(P)$ has $m+k$ vertices.  In other words, 
\[|V(P)| = m+k = |P| -3\;.\]

Conversely, suppose we have an exact subdiagram $(P, V(P))$ of $W = (\cP, n)$ supported on $|V(P)| = |P|+3$ vertices, and let $t$ be the subgraph of $\tau(W)$ corresponding to $(P,V(P))$.

If $|P|=1$, then $(P, V(P))$ is a trivially exact subdiagram. In this case $t$ is an edge internal to $\tau(W)$ connecting two boundary edges, which by definition is a triangulation.

Suppose the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then the corresponding subgraph of $\tau(W)$ has $m-c$ vertices. Call it $t$. If $t$ were a triangulated component, $t$ would have $k -c$ internal edges.

To avoid the issue of exact diagrams being subdiagrams of other exact subdiagrams (for instance, any subdiagram $(q, V_q)$, for $q \in \cP$ is exact), we introduce the notion of maximal exact subdiagrams.

\begin{dfn}
An exact subdiagram $(P, V(P))$ is a {\em maximal exact subdiagram} of $W$ if there is no other exact subdiagram $(Q, V(Q))$ in $W$ that contains $(P,V(P))$ as a strict subdiagram.
\end{dfn}

\begin{cor} \label{uniqueproppartitioncor}
Any admissible Wilson loop diagram $W = (\cP, n)$ can be uniquely decomposed into maximal exact subdiagrams. These maximal subdiagrams partition $\cP$.
\end{cor}

\begin{proof}
Combining Lemmas \ref{decompositionlem} and \ref{lem triang to exact} yields the unique decomposition into maximal exact subdiagrams, and Corollary \ref{maxtriangdisjointcor} ensures that no propagator appears in more than one subdiagram in this decomposition. Since the chord edges of $\tau(W)$ correspond to the propagators of $W$, the decomposition of $\tau(W)$ induces a partition of $\cP$.
\end{proof}

\subsection{Matroid properties of exact subdiagrams}

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely.

\begin{dfn}\label{matroid contraction}
Let $M = (E,\cB)$ be a matroid, and $S \subseteq E$. The {\em contraction} of $M$ by $S$ is the matroid $M/S = (E \setminus S, \cB / S)$, where
\[\cB / S = \{B \setminus S \ \big| \ |B\cap S | \text{ is maximal amonst all }B \in \cB\}.\]
\end{dfn}

In \cite{wilsonloop}, Agarwala and Amat show that certain subdiagrams of $W$ can be realized as contractions of $M(W)$:

\begin{lem} \label{contractsubdiaglem} \cite[Theorem 3.33]{wilsonloop} 
Let $W = (\cP, n)$ be an admissible Wilson loop diagram and $P \subseteq \cP$. If the set $V(P)^c$ has rank $|P^c|$, then the matroid defined by the subdiagram $(P, V(P))$ is equal to the contraction $M(W)/V(P)^c$.
\end{lem}

In Lemma \ref{maxexactcomplementrank} below we show that every exact subdiagram satisfies the rank condition of Lemma \ref{contractsubdiaglem}. In order to do this, we first examine the properties of the set by which we contract.

\begin{dfn}\label{def prop flat}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram and $P \subseteq \cP$. The set $F(P) := V(P^c)^c$ is called the {\em propagator flat} of $P$.
\end{dfn}
Thus in Lemma \ref{contractsubdiaglem} above we would contract $M(W)$ by the propagator flat $F(P^c)$ of the {\em complement} of $P$ in order to study $(P,V(P))$. The justification for this notation is given by (1) of the next lemma: the set $F(P)$ consists of exactly those vertices which {\em only} support propagators in $P$.

\begin{lem}\label{lem properties of prop flats}
Let $F(P)$ be a propagator flat as defined above. Then
\begin{enumerate}
\item $F(P) = V(P) \setminus V(P^c)$.
\item If $Q\subseteq P$ then $F(Q) \subseteq F(P)$.
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem}
\begin{proof}
The proof of (1) and (2) are routine applications of the definition and are omitted.

To prove (3), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ by (1), and no subset of $S$ supports fewer propagators than the number of vertices it contains (this is the definition of an independent set in $M(W)$). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup\{v\}$ also satisfies this independence condition. Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

\begin{lem} \label{lem decompose flat}Let $F$ be a flat in $M(W)$, and let $C \subseteq F$ be the union of all circuits contained in $F$. Then the following are true:
\begin{enumerate}
\item $C = F(\Prop (C))$, i.e. $C$ is a propagator flat.
\item $F \setminus C$ is an independent set. Further, if $W$ does not contain any vertices of rank $0$, i.e. all vertices support at least $1$ propagator, then $F \setminus C$ is an independent flat.
\end{enumerate}
\end{lem}

\begin{proof}
(1) If $F$ is an independent flat, then $C = \emptyset$ and the statement is trivially true. Now suppose that $F$ is a dependent set, so $C$ is non-empty.

Let $v \in C$. Clearly $\Prop(v) \subseteq \Prop(C)$, and so by Lemma \ref{lem properties of prop flats} we have $F(\Prop(v)) \subseteq F(\Prop(C))$. Since $v \in F(\Prop(v))$ by the definition of propagator flat, we have $v \in F(\Prop(C))$ as required.

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. Indeed, by \cite[Proposition 3.10]{wilsonloop}, there exists some $Q \subseteq \Prop(B)$ with $|Q| = |B|$ and such that $B$ is a basis in $M(Q,V(Q))$. We therefore have
\[Q \subseteq \Prop(B) \subseteq \Prop(C),\]
with $|\Prop(C)| = \rk(C)$ by Lemma \ref{lem facts about WLD matroids} and $|Q| = |B| = \rk(C)$ by the choice of $B$. Thus $Q = \Prop(C)$, and so $B$ is a basis in the matroid defined by $W' := (\Prop(C),V(\Prop(C)))$. All that remains is to note that $w \in F(\Prop(C)) \subseteq V(\Prop(C))$, i.e. $w$ is a vertex in $W'$; thus $B \cup\{w\}$ is dependent in $W'$, and hence also in $W$.

Returning our attention to $M(W)$, the dependent set $B \cup \{w\}$ must contain a circuit $C'$, with $w \in C'$ since $B$ was an independent set. Now $C \cup C' = C \cup \{w\} \supsetneq C$ is a cycle, \hlfix{which contradicts the maximality of $C$}{as long as $w \in F$. I can't figure out why this has to be true?}

For part (2), first note that $F \setminus C$ is automatically independent as it contains no circuits. Now suppose that all vertices in $W$ have rank $\geq 1$.

For any $e \not\in F$, we certainly have $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C) +1$ since $F$ is a flat. Now let $e \in C$, and suppose that $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C)$. This implies that $(F\setminus C)\cup \{e\}$ is dependent, and hence contains a circuit. This circuit must contain at least two elements (since all vertices support at least one propagator), but this contradicts the fact that $C$ was the union of all circuits in $F$.

Thus $\rk((F\setminus C)\cup \{e\}) = \rk(F\setminus C) + 1$ for any $e \not\in F\setminus C$, and hence $F\setminus C$ is a flat.
\end{proof}

\begin{cor} \label{classifyflats}
If $F$ is a flat of a Wilson loop diagram, it can be written as the disjoint union of a cyclic propagator flat and an independent flat. \end{cor}

@@

%Finally, we note the differences between subdiagrams of Wilson loop diagrams, and restrictions or cotractions of the associated matroids.

%Given any matroid, one may restrict it to a subset of the base set. The bases of the restricted matroid come from intersecting bases of the original with the subset. It is worth noting that a the matroid defined by a subdiagram is different from the restriction of the matroid of a Wilson loop diagram to a set of vertices.

% \begin{dfn} \label{restrictiondfn}
% For $W = (\cP, n)$, the restricted diagram, $W|_V$ is the matroid defined by only looking at the vertices $V \subset [n]$.
% \end{dfn}

% The key difference between a subdiagram and a restriction is that the propagator support function does not change in the case of restriction, while it may in the case of a subdiagram. In particular, for $v \in V$, $\Prop_W(v) = \Prop_{W|_V}(v)$, while $\Prop_{(P, V(P))} (v) = \Prop_W(v) \cap P$.

% A subdiagram is more closely related to a contracted matroid.

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, n)$ be a Wilson loop diagram, and $P \subsetneq \cP$. Then:
\begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.
\end{enumerate}
\end{lem}

\begin{proof}
We first verify that $F(P^c) = V(P)^c$ is non-empty and dependent for any exact subdiagram $(P,V(P))$. The definition of exactness means that the vertices $V(P)$ cannot support a larger set of propagators (otherwise $W$ would no longer be admissible). Therefore, since $P \subsetneq \cP$, there must be at least one propagator supported on $V(P)^c$, i.e. it cannot be empty.

Since $W$ is admissible, we have $n \geq |\cP| + 4$. Rewriting this as
\[|V(P)| + |F(P^c)|  \geq  |P| + |P^c| +4,\]
and combining it with the fact that $|V(P)| = |P| + 3$ (from the exactness of $(P,V(P))$), we obtain
\begin{equation}\label{eq F is dependent}|F(P^c)| > |P^c| \;.\end{equation}
By Lemma \ref{lem properties of prop flats}, $F(P^c)$ is the set of vertices that {\em only} supports propagators in $P^c$, so in particular we have $\Prop(F(P^c)) = P^c$. Equation \eqref{eq F is dependent} is therefore saying that $F(P^c)$ supports fewer propagators than the number of vertices it contains, i.e. $F(P^c)$ is dependent.

We address part (2) first. Suppose that $(P, V(P))$ is a maximal exact subdiagram; by Corollary \ref{classifyflats} we can decompose $F(P^c)$ as
\begin{equation}\label{eq decompose F}F(P^c) = C \sqcup S,\end{equation}
where $C$ is the largest cyclic flat contained in $F(P^c)$ and $S$ is an  independent set. Note that $C$ must be non-empty since $F(P^c)$ is dependent.
